The Effect of Elections and Changed Governments on Discussion in the Australian Federal Parliament (1901--2017)

### 2018-11-19 outputs/docs/events/paper.Rmd
Update paper

# Introduction
Government policy is partly driven by parliamentary discussion. Conversely, that same discussion can indicate a government's priorities. But major events -- be they expected, such as an election, and unexpected, such as a recession or terrorist attack -- can affect the course of parliamentary discussion. For instance, think of how a new government often goes to some trouble to seem different to the one they replace, or how events such as the 9/11 attacks altered government priorities.

In this paper we examine text records of what was said in the Australian Federal Parliament. We create a dataset based on all sitting days from 1901 through to 2017. We have records for 7,869 days in the House of Representatives and 6,682 days in the Senate. Our dataset consists of counts of every word that occurs at least thrice, after we convert all words to lower-case; remove names, numbers, and punctuation; join commonly co-located words; and remove stop words.

@@

Our work contributes to a growing modern quantitative social sciences literature that considers text as an input to statistical methods, rather than requiring separate analysis. This literature sits across and draws from various historically-separate disciplines including applied statistics, economics, and political science. We contribute to this literature in terms of both data and methods. From a data perspective we bring to bear an essentially-complete record of what was said in the Australian Federal Parliament on a daily basis, and we make our dataset available to other researchers via the R package \texttt{hansardr}. From a methods perspective, our analysis model has several advantages over existing methods. These include: allowing the events to have more-complicated auto-correlated functional forms; implementing pooling across groups of similar documents; and identifying outlying topic distributions without the need to pre-specify the event of interest. As the digitisation of historical sources continues and computational power becomes cheaper, we expect interest in a quantitative approach to text to only increase and the data and methods in our paper will be extended in a variety of ways.

@@

Following the example of the UK a daily text record called Hansard of what was said in the Australian Federal Parliament has been made available since it was established in 1901.[^Verbatimfootnote] Earlier work on the influence of parliaments, such as @ZandenBuringhBosker2012, often examined broader activity measures such as counts of sitting days. This allowed for long time-frames and wide comparisons. But analysing Hansard records and their equivalents directly is increasingly viable as new methods and reduced computational costs make it easier.

@@

The recent digitisation of the Canadian Hansard, @BeelenEtc2017, has allowed text analysis of Canadian parliament records. For instance, @RheaultCochran2018 examined ideology and party polarisation in Britain and Canada. A digitised version of the UK Hansard has been available for some time and there has been much text analysis of it. For instance, @Duthie2016 examined which politicians made supportive or aggressive statements toward other politicians and @PetersonSpirling2018 examined polarisation. One exciting aspect of research using the UK Hansard has been linking text records with other datasets. For instance, @SlapinKirklandLazzaroLeslie2018 linked votes and speeches to examine grandstanding within parties. Another exciting aspect is that as digitisation methods improve, increasingly older UK records can be analysed, for instance, @Dimitruk2018 considered the effect of estate bills on prorogations in seventeenth century England. In New Zealand, @Curran2018 modelled the topics discussed between 2003 and 2016, and @Graham2016 examined unparliamentary language between 1890 and 1950. And in the US, @GentzkowShapiroTaddy2018 examined congressional speech records from 1873 to 2016 to find that partisanship has risen in the past few decades.

Parts of Australian Hansard records have been analysed for various purposes. For instance, @Rasiah2010 examined Hansard records for the Australian House of Representatives to examine whether politicians attempted to evade questions about Iraq during February and March 2003. @GansLeigh2012 examined Australian Hansard records to associate mentions by politicians of certain public intellectuals with neutral or positive sentiment. And @Salisbury2011 examined unparliamentary behaviour. @FraussenGrahamHalpin2019 examined Australian Hansard records to assess the prominence of interest groups. The closest research to ours that we have found is @Boulous2013 who examined parliamentary debate in Australia for the period 1946 to 2012.

The Australian Federal Parliament makes daily Hansard records available online as PDFs and these are considered the official release. Additionally, XML records are available in some cases.[^XMLfootnote] We provide an example of a Hansard PDF page in Appendix \@ref(examplehansardpage). There are 14,551 days of publicly available Hansard records across both chambers of the Australian Federal Parliament that we have PDFs for and further summary statistics for this are provided in Appendix \@ref(hansardsummarystatistics). Our data cleaning process indicates concerns with a small number of PDFs and these are detailed in Appendix \@ref(knownhansardissues).

[^XMLfootnote]: Tim Sherratt makes these XML records available as a single download and also presents them in a website -- http://historichansard.net/ -- that can be used to explore Commonwealth Hansard records from 1901 to 1980. Commonwealth XML records from 1998 to 2014 are available from Andrew Turpin’s website, and from 2006 through to today from Open Australia’s website. The records can also be downloaded from the Australian Hansard website.

@@

The goal of our modelling strategy is twofold. Firstly, we want to use topic modelling [@Blei2003latent] to summarise the Hansard text into meaningful topics that reduce the dimensionality of the text data and capture the main themes discussed in parliament over time. Secondly, we want to relate the resulting topic distributions to temporal trends, changes, and events, such as a change in government, elections, and other external events.

We first use a Correlated Topic Model (CTM) [@BleiLafferty2007] to obtain estimated topic distributions over time. We consider these topic distributions as inputs that can be analysed within another model. Thus, the second modelling step involves using a Bayesian hierarchical Dirichlet model to analyse changes in the topic distributions (obtained from the first step) in relation to events of interest.

@@

Other work such as @BaumgartnerJones1993 and @DowdingHindmoor2010 addresses this problem by creating a standardised codebook of policy categories and sub-categories and then manually assigning text to topics as appropriate. This approach ensures the categorisation is reasonable but as it is a manual process the size of the text that can be categorised is limited.

In order to effectively categorise topics over the whole period covering Hansard, we chose to use topic modelling , a statistical technique which aims to extract the underlying or latent 'topics' from a collection of texts. In particular, we use a method which is similar to Latent Dirichlet Allocation (LDA), first developed by @Blei2003latent.

The key assumption behind LDA is that each day's text, 'a document', in Hansard is made by speakers who decide the topics they would like to talk about in that document, and then choose words, 'terms', that are appropriate to those topics. A topic could be thought of as a collection of terms, and a document as a collection of topics, where these collections are defined by probability distributions. The topics are not specified *ex ante*; they are an outcome of the method, and it is in this sense that this approach can be considered unsupervised machine learning. Terms are not necessarily unique to a particular topic, and a document could be about more than one topic. The goal is to have the words found in each day's Hansard group themselves to define topics. This can provide more flexibility than other approaches such as a strict word count method, but can require a larger dataset and make interpretation more difficult. More detail on how topic modelling works is available in Appendix \@ref(LDAexample).

@@

As such, we chose to use a Correlated Topic Model (CTM) [@BleiLafferty2007] to obtain topic distributions. CTM is a slight extension to LDA which allows for correlations between topics. More details about CTM can be found in Appendix \@ref(LDAexample).

@@

One advantage of our model over using the STM is that we can restrict the effect of an election to be monotonically decreasing. This allows us to identify differences between government and election effects even when there is a one-term government. The value of the initial effect, $\alpha_{e[s],0,p}$, and the AR(1) term, $\rho$, both have non-informative priors:
$$
\alpha_{e[s],0,p} \sim \mbox{Normal}(0, 100);\mbox{ and}\\
\rho_{e[s],p} \sim \mbox{Uniform}(0,1).
$$

We model the underlying time trend in topics using splines regression. The intuition behind this term is to capture the underlying non-linear trend in topic distributions over time, which is caused by large-scale structural changes in the economy, and Australian society and culture. The $x_{s,k}$ for $k = 1,2,\dots, K$ are the value of cubic basis splines for sitting period $s$ at knot point $k$. We place knot points every five sitting periods as this is the average length of time for a government to sit. Non-informative priors are placed on the splines coefficients:
$$
\beta_{p,k} \sim \mbox{Normal}(0, 100).
$$

Finally, the sitting period-specific random effect $\delta_{s,p}$ allows for the topic distributions in some sitting periods to be different than expected based on government and election effects. This allows us to identify large deviations away from the expected distribution, thus helping to identify the effect of other, non-government and non-election events. In addition, this set up also partially pools effects across sitting periods. The $\delta_{s,p}$ values are modelled as:
$$
\delta_{s,p} \sim \mbox{Normal}(0, \sigma_{e[s],p}^2).
$$

@@

Figure \@ref(fig:exampletopics) illustrates the CTM output based on a sample from the Australian House of Representatives between 1901 and 2017. It shows how each day's parliamentary discussion can be apportioned to a topic and highlights how these proportions change over time. The highlighted topics are those related to war and conflict. As expected, the two chambers of the Australian Federal Parliament are fairly similar.

@@

We summarise our results in terms of governments in Table \@ref(tab:governmentsresults) and in terms of elections in Table \@ref(tab:electionsresults). These tables focus on elections and governments that were different to the ones that preceded them. Complete lists of elections and governments are available in Appendices \@ref(fulllistofelections) and \@ref(fulllistofgovernments).

elections_results %>%
  kable(booktabs = T, caption = "Significant elections, and neighours") %>%
  kable_styling(font_size = 10)

@@

@@

Once government and election effects have been taken into consideration, some days stand out compared with others in their sitting period. We do not explicitly include them in the model because of over-fitting and effect-type concerns, but we are interested to see if these can be explained by events that occurred on or before that sitting day. For instance, we may expect that events of a historical magnitude, such as the 9/11 attacks would be the focus of an entire day's discussion, or that the sitting day when the Apology to the Stolen Generation was delivered, or some particular legislation introduced, would be different to others in that sitting period.

@@

# Discussion
Of the 36 governments over this period, we find that 20 of them are significantly different to the government that preceded them. However, four of these results -- the significance of the Page Government, the First Menzies Government, the Chifley Government and the Abbott Government -- are likely due to the short length of either the government or its predecessor and should be ignored, leaving only 16. The earliest of these is the First Fisher Government, the Cook Government and the Third Fisher Government, which were different to the Second Deakin Government, the Second Fisher Government, and the Cook Government, respectively. To a certain extent this is likely due to changes due to World War I.

The Second Menzies Government, starting in 1949, is the next government that is significantly different to a predecessor. The rest of the governments that are different are concentrated in the second half of our sample, with three of them being in the past twenty years. Similarly, of the 45 general elections that have been held we find that 14 of them define periods that were significantly different to their immediate predecessor. 1974, 1980, 1990, 1998, 2004, and 2007 stand out as elections where the government did not change, but the model suggests there was considerable change in the topics discussed in parliament.

@@

@@

## Overview and example
As applied to Hansard, LDA considers each statement to be a result of a process where a politician first chooses the topics they want to speak about. After choosing the topics, the politician then chooses appropriate words to use for each of those topics. Statistically, LDA considers each document as having been generated by some probability distribution over topics. Similarly, each topic is considered a probability distribution over terms. To choose the terms used in each document, terms are picked from each topic in the appropriate proportion.

>>>>

# Introduction
Government policy is partly driven by parliamentary discussion. Conversely, that same discussion can indicate a government's priorities. But major events -- be they somewhat expected, such as an election, and unexpected, such as a recession or terrorist attack -- can affect the course of parliamentary discussion. For instance, think of how a new government often goes to some trouble to seem different to the one they replace, or how events such as the 9/11 attacks altered government priorities.

In this paper we examine text records of what was said in the Australian Federal Parliament. We create a dataset based on the sitting days from 1901 through to 2017. We have records for 7,869 days in the House of Representatives and 6,682 days in the Senate. Our dataset consists of counts of every word that occurs at least thrice, after we convert all words to lower-case; remove names, numbers, and punctuation; join commonly co-located words; and remove stop words.

@@

Our work contributes to the growing literature in quantitative social science that considers text as an input to statistical methods, rather than requiring separate analysis. This literature sits across and draws from various historically-separate disciplines including applied statistics, economics, and political science. We contribute to this literature in terms of both data and methods. From a data perspective we bring to bear an essentially-complete record of what was said in the Australian Federal Parliament on a daily basis, and we make our dataset available to other researchers via the R package \texttt{hansardr}. From a methods perspective, our analysis model has several advantages over existing methods. These include: allowing the events to have more-complicated auto-correlated functional forms; implementing pooling across groups of similar documents; and identifying outlying topic distributions without the need to pre-specify the event of interest. As the digitisation of historical sources continues and computational power becomes cheaper, we expect the popularlity of quantitative approaches to text to increase further and the data and methods in our paper to be extended in a variety of ways.

@@

Following the example of the UK a daily text record called Hansard of what was said in the Australian Federal Parliament has been made available since Federation in 1901.[^Verbatimfootnote] Analysing Hansard records and their equivalents directly is increasingly viable as new methods and reduced computational costs make it easier.

@@

The recent digitisation of the Canadian Hansard, @BeelenEtc2017, has allowed increased analysis of Canadian parliament text records. For instance, @RheaultCochran2018 examined ideology and party polarisation in Britain and Canada. A digitised version of the UK Hansard has been available for some time. For instance, @Duthie2016 examined which politicians made supportive or aggressive statements toward other politicians, and @PetersonSpirling2018 examined polarisation. One exciting aspect of research using the UK Hansard has been linking text records with other datasets. For instance, @SlapinKirklandLazzaroLeslie2018 linked votes and speeches to examine grandstanding within parties. Another exciting aspect is that as digitisation methods improve, increasingly older UK records can be analysed, for instance, @Dimitruk2018 considered the effect of estate bills on prorogations in seventeenth century England. In New Zealand, @Curran2018 modelled the topics discussed between 2003 and 2016, and @Graham2016 examined unparliamentary language between 1890 and 1950. And in the US, @GentzkowShapiroTaddy2018 examined congressional speech records from 1873 to 2016 to find that partisanship has risen in the past few decades.

Parts of Australian Hansard records have been analysed for various purposes. For instance, @Rasiah2010 examined Hansard records for the Australian House of Representatives to examine whether politicians attempted to evade questions about Iraq during February and March 2003. @GansLeigh2012 examined Australian Hansard records to associate mentions by politicians of certain public intellectuals with neutral or positive sentiment. @Salisbury2011 examined unparliamentary behaviour. And @FraussenGrahamHalpin2019 examined Australian Hansard records to assess the prominence of interest groups. The closest research to ours that we have found is @Boulous2013 who examined parliamentary debate in Australia between 1946 and 2012.

The Australian Federal Parliament makes daily Hansard records available online as PDFs and these are considered the official release. Although we do not use them here, XML records are also available in most cases.[^XMLfootnote] We provide an example of a Hansard PDF page in Appendix \@ref(examplehansardpage). There are 14,551 days of publicly available Hansard records across the two chambers of the Australian Federal Parliament for which we have PDFs and further summary statistics for this are provided in Appendix \@ref(hansardsummarystatistics). Our data cleaning process indicates concerns with a small number of PDFs and these are detailed in Appendix \@ref(knownhansardissues).

[^XMLfootnote]: Tim Sherratt makes Commonwealth XML records for 1901 to 1980 available as a single download at: http://historichansard.net/. Commonwealth XML records from 1998 to 2014 are available from Andrew Turpin’s website, and from 2006 through to today from Open Australia’s website. The records can also be downloaded from the Australian Hansard website or the website can be scraped.

@@

The goal of our modelling strategy is twofold. Firstly, we want to use topic modelling [@Blei2003latent] to summarise the Hansard text into meaningful topics that reduce the dimensionality of the text data and capture the main themes discussed in parliament over time. Secondly, we want to relate the resulting topic distributions to temporal trends, changes, and events, such as a change in government, elections, and other external events.  The output of the first stage is the input to the second stage.

We first use a Correlated Topic Model (CTM) [@BleiLafferty2007] to obtain estimated topic distributions over time. We consider these topic distributions as inputs that can be analysed by another model. Thus, the second modelling step involves using a Bayesian hierarchical Dirichlet model to analyse changes in the topic distributions (obtained from the first step) in relation to events of interest.

@@

Other work such as @BaumgartnerJones1993 and @DowdingHindmoor2010 addressed this problem by creating a standardised codebook of policy categories and sub-categories and then manually assigning text to topics as appropriate. This approach ensures the categorisation is reasonable but as it is a manual process the size of the text that can be categorised is limited.

In order to effectively categorise the topics of the entire Hansard, we chose to use topic modelling, a statistical technique which aims to extract the underlying or latent 'topics' from a collection of texts. In particular, we use a method which is similar to Latent Dirichlet Allocation (LDA), first developed by @Blei2003latent.

The key assumption behind LDA is that each day's text, 'a document', in Hansard is made by speakers who decide the topics they would like to talk about in that document, and then choose words, 'terms', that are appropriate to those topics. A topic could be thought of as a collection of terms, and a document as a collection of topics, where these collections are defined by probability distributions. The topics are not specified *ex ante*; they are an outcome of the method, and it is in this sense that this approach can be considered unsupervised machine learning. Terms are not necessarily unique to a particular topic, and a document could be about more than one topic. The goal is to have the words found in each day's Hansard group themselves to define topics. This can provide more flexibility than other approaches such as a strict word count method, but can require a larger dataset and make interpretation more difficult. A

An overview and simplified example of how topic modelling works is available in Appendix \@ref(LDAoverviewandexample). The underlying document generation process is discussed in Appendix \@ref(LDAdocgenprocess), and Appendix \@ref(LDAposteriorestimation) explains how this is reversed to generate topics. Finally, the choice of the number of topics has a large effect on the composition of the topics, and some diagnostics for selecting the appropriate number are discussed in Appendix \@ref(selecttopicnumber).

@@

As such, we chose to use a Correlated Topic Model (CTM) [@BleiLafferty2007] to obtain topic distributions. CTM is a slight extension to LDA which allows for correlations between topics. More details about CTM can be found in Appendix \@ref(correlatedtopicmodelsection).

@@

One advantage of our model over using the STM is that we can restrict the effect of an election to be monotonically decreasing. This allows us to identify differences between government and election effects even when there is a one-term government. The value of the initial effect, $\alpha_{e[s],0,p}$, and the AR(1) term, $\rho$, both have non-informative priors:
$$
\begin{aligned}
\alpha_{e[s],0,p} &\sim \mbox{Normal}(0, 100) \mbox{ and} \nonumber \\
\rho_{e[s],p} &\sim \mbox{Uniform}(0,1). \nonumber
\end{aligned}
$$

We model the underlying time trend in topics using splines regression. The intuition behind this term is to capture the underlying non-linear trend in topic distributions over time, which is caused by large-scale structural changes to Australian society, culture, and the economy. The $x_{s,k}$ for $k = 1,2,\dots, K$ are the value of cubic basis splines for sitting period $s$ at knot point $k$. We place knot points every five sitting periods as this is the average length of time for a government to sit. Non-informative priors are placed on the splines coefficients:
$$
\beta_{p,k} \sim \mbox{Normal}(0, 100).
$$

Finally, the sitting-period-specific random effect $\delta_{s,p}$ allows for the topic distributions in some sitting periods to be different than expected based on government and election effects. This allows us to identify large deviations away from the expected distribution, thus helping to identify the effect of other, non-government and non-election events. In addition, this set up also partially pools effects across sitting periods. The $\delta_{s,p}$ values are modelled as:
$$
\delta_{s,p} \sim \mbox{Normal}(0, \sigma_{e[s],p}^2).
$$

@@

Figure \@ref(fig:exampletopics) illustrates the CTM output based on a sample from the Hansard. It shows how each day's parliamentary discussion can be apportioned to a topic and highlights how these proportions change over time. The highlighted topics are those related to war and conflict. As expected, the two chambers of the Australian Federal Parliament are fairly similar.

@@

We summarise our results in terms of governments in Table \@ref(tab:governmentsresults) and in terms of elections in Table \@ref(tab:electionsresults). These tables focus on elections and governments that were different to the ones that preceded them. Complete lists of the Australian elections and governments are available in Appendices \@ref(fulllistofelections) and \@ref(fulllistofgovernments).

@@

elections_results %>%
  kable(booktabs = T, caption = "Elections that were significantly different to the one before") %>%
  kable_styling(font_size = 10) %>% 
  footnote(general = "Here is a general comments of the table. ")

@@

Once government and election effects have been taken into consideration, some days stand out compared with others in their sitting period. We do not explicitly include them in the model because of over-fitting and effect-type concerns, but we are interested to see if these can be explained by events that occurred on or before that sitting day. For instance, we may expect that events of a historical magnitude, such as the 9/11 attacks would change the discussion, or that the sitting day when, for instance, the Apology to the Stolen Generation was delivered, or some particularly prominent legislation introduced, would be different to others in that sitting period.

@@

# Discussion
Of the 36 governments over this period, we find that 20 of them are significantly different to the government that preceded them. However, four of these results -- the significance of the Page Government, the First Menzies Government, the Chifley Government and the Abbott Government -- are likely due to the short length of either the government or its predecessor and should be ignored, leaving only 16. The earliest of these are the First Fisher Government, the Cook Government and the Third Fisher Government, which were different to the Second Deakin Government, the Second Fisher Government, and the Cook Government, respectively. To a certain extent this is likely due to changes due to World War I.

The Second Menzies Government, begining in 1949, is the next government that is significantly different to its predecessor. The other governments that are different are concentrated in the second half of our sample, with three of them being in the past twenty years. Similarly, of the 45 general elections that have been held we find that 14 of them define periods that were significantly different to their immediate predecessor. 1974, 1980, 1990, 1998, 2004, and 2007 stand out as elections where the government did not change, but the model suggests there was considerable change in the topics discussed in parliament.

@@

In this paper we think of events as affecting discussion in parliament. Given the events that we consider and the broad topics of discussion we consider this the most appropriate direction for this paper. Other papers could consider an alternative specification, especially if they were to focus on particular aspects.

@@

## Overview and example {#LDAoverviewandexample}
As applied to Hansard, LDA considers each statement to be a result of a process where a politician first chooses the topics they want to speak about. After choosing the topics, the politician then chooses appropriate words to use for each of those topics. Statistically, LDA considers each document as having been generated by some probability distribution over topics. Similarly, each topic is considered a probability distribution over terms. To choose the terms used in each document, terms are picked from each topic in the appropriate proportion.
